---
title: "HW5 Justin Kim jtk2141"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goals: Simulating probability distributions using the Inverse Transform Method and AcceptReject Method. Built in R distribution functions. Estimate the mathematical constant π
using Monte Carlo techniques. More practice writing functions and plotting.
First set your seed as 0, i.e., set.seed(0)

Part 1: Inverse Transform Method

1. Let U be a uniform random variable over [0,1]. Find a transformation of U that allows
you to simulate X from U.

```{r}
finv <- function(u) {
  return(ifelse((u<0|u>1),0,tan(pi*(u-1/2))))
}
```

2. Write a R function called cauchy.sim that generates n simulated Cauchy random
variables. The function should have the single input n and should use the inversetransformation from Part 1. Test your function using 10 draws.

```{r}
set.seed(0)
cauchy.sim <-function(n) {
  u <-runif(n)
  return(ifelse((u<0|u>1),0,tan(pi*(u-1/2))))
}
cauchy.sim(10)
```
3. Using your function cauchy.sim, simulate 1000 random draws from a Cauchy distribution. Store the 1000 draws in the vector cauchy.draws. Construct a histogram of the simulated Cauchy random variable with fX(x) overlaid on the graph. Note: when plotting the density curve over the histogram, include the argument prob = T. Also
note: the Cauchy distribution produces extreme outliers. I recommend plotting the histogram over the interval (−10, 10)

```{r}
set.seed(0)
cauchy.draws <-cauchy.sim(1000)
hist(cauchy.draws,prob =T,breaks=2000,xlim=(c(-10,10)))
x <-seq(-10,10,0.01)
lines(x,1/(pi*(1+x^2)),col ="red")

```

Part 2. Rejeft-Accept Method

Part 2

4. Write a function f that takes as input a vector x and returns a vector of f(x) values. Plot the functionbetween -3 and 3. Make sure your plot is labeled appropriately.
```{r}
f <-function(x){
  return(ifelse((x< -1 | x>2),0,1/9*(4-x^2)))
}
x <-seq(-3,3,0.01)
plot(x,f(x),type="l",ylab="f(x)",main="PDF")
```
5. Determine the maximum of f(x) and find an envelope function e(x) by using a uniform
density for g(x). Write a function e which takes as input a vector x and returns a vector
of e(x) values

maximum: 0.44444
```{r}
maximumofx <- 0
maximumoffx <- f(maximumofx)
e <- function(x) {
  return(ifelse((x < -1 | x > 2), Inf, maximumoffx))
}
maximumoffx
```
6. Using the Accept-Reject Algorithm, write a program that simulates 10,000 draws
from the probability density function f(x) from Equation 1. Store your draws in the
vector f.draws.
```{r}
set.seed(0)
m <- 0
pdraws <- 10000
f.draws <- numeric(pdraws)
while (m < pdraws) {
  y <- runif(1, -1, 2)
  u <- runif(1)
  if (u < f(y)/e(y)) {
    m <- m + 1
    f.draws[m] <- y
    }
}
```
7. Plot a histogram of your simulated data with the density function f overlaid in the
graph. Label your plot appropriately.
```{r}

hist(f.draws, prob = T, ylab="f(x)", xlab = "x", main="Accept-Reject
Algorithm", xlim=(c(-3, 3)))
x <- seq(-3, 3, 0.01)
lines(x, f(x), col="red")
```
Problem 3: Reject-Accept Method Continued

8. Write a function f that takes as input a vector x and returns a vector of f(x) values.
Plot the function between −5 and 5. Make sure your plot is labeled appropriately
```{r}
x <- seq(-5, 5, 0.01)
f <- function(x) {
return(1/sqrt(2*pi)*exp(-1/2*x^2))
}
plot(x, f(x), type="l", ylab="f(x)", main="Standard Normal Distribution", col= "red")

```

9. Write a function e that takes as input a vector x and constant alpha (0 < α < 1)
and returns a vector of e(x) values. The envelope function should be defined as
e(x) = g(x)/α.
```{r}
e <- function(x, alpha) {
  if(alpha < 1 && alpha >0){
return((1/pi)*(1/(1+x^2))*(1/alpha))
  }
  return((1/pi)*(1/(1+x^2))*(1/alpha))
}
```
10. Determine a “good” value of α. You can solve this problem graphically. To show your
solution, plot both f(x) and e(x) on the interval [−10, 10].

A good alpha value would be 0.8
```{r}
x <- seq(-10, 10, 0.01)
plot(x, f(x), type="l", col = "red")
lines(x, e(x, alpha=0.8), col="blue")
```
11. Write a function named normal.sim that simulates n standard normal random variables using the Accept-Reject Algorithm. The function should also use the InverseTransformation from Part 1. Test your function using n=10 draw.
```{r}
   normal.sim <- function(x){
     n<-0
     f.draws <- numeric(x)
     while(n < x){
       y <- rnorm(1)
       u <- rnorm(1)
       if (u < finv(y)/e(y, .1)){
         n <- n+1
         f.draws[n] <- y
       }
     }
     return(f.draws)
   }

normal.sim(10)
```
12. Using your function normal.sim, simulate 10,000 random draws from a standard
normal distribution. Store the 10,000 draws in the vector normal.draws. Construct

a histogram of the simulated standard normal random variable with f(x) overlaid on
the graph. Note: when plotting the density curve over the histogram, include the
argument prob = T
```{r}

normal.draws <- normal.sim(10000)
hist(normal.draws, prob = T, ylab="f(x)", main="Standard Normal Distribution", col= "red")
lines(x, f(x), lty = 1)
```
13. Write a while() loop to implement this procedure. Importantly, save all the positive
values of x that were visited in this procedure in a vector called x.vals, and display
its entries
```{r}
set.seed(0)

x <- 5

x.vals <- c(x)

while (x > 0) {
  r <- runif(1, min=-2, max=1)
  x <- x + r
  if (x > 0) {
    x.vals <- c(x.vals, x)
    }
  }
x.vals
```

14. Produce a plot of the random walk values x.vals from above versus the iteration
number. Make sure the plot has an appropriately labeled x-axis and and y-axis. Also
use type="o" so that we can see both points and lines.

```{r}
plot(x.vals, type="o", xlab = "Iteration Number", ylab="X.Vals")

```

15. Write a function random.walk() to perform the random walk procedure that you
implemented in question (9). Its inputs should be: x.start, a numeric value at which
we will start the random walk, which takes a default value of 5; and plot.walk, a
boolean value, indicating whether or not we want to produce a plot of the random
walk values x.vals versus the iteration number as a side effect, which takes a default
value of TRUE. The output of your function should be a list with elements: x.vals,
a vector of the random walk values as computed above; and num.steps, the number
of steps taken by the random walk before terminating. Run your function twice with
the default inputs, and then twice times with x.start equal to 10 and plot.walk =
FALSE.

```{r}
set.seed(0)
random.walk <- function(x.start=5, plot.walk=TRUE) {
  numbersteps <- 0
  x <- x.start
  x.vals <- c(x)
  while (x > 0) {
    x <- x + r
    r <- runif(1, min=-2, max=1)
    if (x > 0) {
      x.vals <- c(x.vals, x)
      num.steps <- numbersteps + 1
    }
    }
  if (plot.walk == TRUE) {
    plot(x.vals, type="o", xlab = "Iternation Number", ylab="X.Vals")
    }
  return(list(x.vals, num.steps))
  }
random.walk(5)
random.walk(5)
random.walk(10, plot.walk = FALSE)
```

16. We’d like to answer the following question using simulation: if we start our random
walk process, as defined above, at x = 5, what is the expected number of iterations
we need until it terminates? To estimate the solution produce 10, 000 such random
walks and calculate the average number of iterations in the 10, 000 random walks you
produce. You’ll want to turn the plot off here.

```{r}
set.seed(0)
iteration <- 0
n <- 0
while (n < 10000) {
  steps <- random.walk(5, plot.walk=FALSE)[[2]]
  iteration <- iteration + steps
  n <- n + 1
}
mean(iteration)
```

17.  Modify your function random.walk() defined previously so that it takes an additional
argument seed: this is an integer that should be used to set the seed of the random number generator, before the random walk begins, with set.seed(). But, if
seed is NULL, the default, then no seed should be set. Run your modified function
random.walk() function twice with the default inputs, then run it twice with the
input seed equal to (say) 33 and plot.walk = FALSE
```{r}

set.seed(0)
random.walk <- function(x.start=5, plot.walk=TRUE, seed = NULL) {
  if(!is.null(seed)){
    set.seed(seed)
  }
  numbersteps <- 0
  x <- x.start
  x.vals <- c(x)
  while (x > 0) {
    x <- x + r
    r <- runif(1, min=-2, max=1)
    if (x > 0) {
      x.vals <- c(x.vals, x)
      num.steps <- numbersteps + 1
    }
    }
  if (plot.walk == TRUE) {
    plot(x.vals, type="o", xlab = "Iternation Number", ylab="X.Vals")
    }
  return(list(x.vals, num.steps))
}
random.walk(plot.walk = FALSE, seed = 33)
random.walk(plot.walk = FALSE, seed = 33)
```

18. Run the following code:
```{r}
g <- function(x) {
return(exp(-x^3))
}
x <- seq(0,1,.01)
alpha <- 2
beta <- 2
plot(x,g(x),type="l",xlab="x",ylab="",ylim=c(-.1,1.4))
polygon(c(0,seq(0,1,0.01),1),c(0,g(seq(0,1,0.01)),0) ,col="pink")
lines(x,rep(1,length(x)),col="red")
lines(x,dbeta(x,shape1=alpha,shape2=beta),col="blue")
legend("topleft",legend=c("g(x)","uniform","beta(2,2)"),
       lty=c(1,1,1),col=c("black","red","blue"),cex=.6)
```
19:
```{r}
n <- 1000^2
mean(g(runif(n)))
```
20:
```{r}
mean(g(rbeta(n, shape1 = alpha, shape2 = beta)))
```

